---
title: "split_drop"
author: "Todd Farmer"
date: "09/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# This document show how the Drop-Seq dataset used in the Murai lab is contructed from the provided data.
```{r}
library(R.matlab)
library(Seurat)
library(dplyr)
library(DropSeq.util)
library(gsubfn)
library(sctransform)
library(future)
# plan("multiprocess", workers = 4)
options(future.globals.maxSize = 10000 * 1024^2)
source("DropSeuratFunctions.R")
```
# The Data
The data is split in a digital gene expression (DGE) matrix for each brain region, an annotation file for each region, and a annotation file that describes all of the cell types annotated in the publication.

## DGE
The expresssion matrix is a sparse matrix that consists of cells in columns and genes in rows
Here is small subset of a DGE:
## Outcomes

## Annotation


## Set the directories to use
```{r}

path <- "/Volumes/one/DATA/drop-seq/raw/"

files <- read.csv("/Volumes/one/DATA/drop-seq/raw/dropSeq_files.csv")
```

## Download data
DGE matrix for each region
as well as metadata(subcluster assignment etc)
```{r download}

```


## Construct seurat object for each brain region

```{r extractData}
path <- "raw/"
idents <- readRDS(paste0(path, "annotation.BrainCellAtlas_Saunders_version_2018.04.01.RDS"))

files <- read.csv(paste0(path, "dropSeq_files.csv"), header = FALSE)
for (row in 1:nrow(files)) {
    rawDGE <- paste0(path, as.character(files[row, 1]))
    RDS <- paste0(path, as.character(files[row, 2]))
    prefix <- as.character(files[row, 3])
    print(paste0("Contructing ", prefix, " using ", rawDGE, " and ", RDS))
    SO <- extractDGE(rawDGE, RDS, idents, prefix)
    print(SO)
    }
    

```

# Calculate each brain region
```{r}

# loop over files in folder
SOs <- list.files('annotated/', pattern='*\\_annotated.RDS', recursive=FALSE, full.names=TRUE)

for (file in SOs){
  SO <- readRDS(file)
  print(file)
  print(SO)
  SO <- dropPass(SO)
  SO <- calcDrop(SO)
  message("Writing to disk...")
  saveRDS(SO, paste0("TISSUE/", gsub("_annotated.RDS","", basename(file)), ".RDS"))
}
```


## Generate an Seurat objec that contains all cells.
```{r combineSOs}
annotated_dir <- "annotated/"
SOs <- list.files(path = "annotated/")
AllDrop <- readRDS(paste0(annotated_dir, SOs[1]))

for (SO in 2:length((SOs))){
  print(SOs[SO])
  SO <- readRDS(paste0(annotated_dir, SOs[SO]))
  AllDrop <- Seurat:::merge.Seurat(x = AllDrop, y = SO)
}
saveRDS(AllDrop, file="annotated/ALL_annotated.RDS")
AllDrop
```




## Split by Class


```{r}
# ALL <- readRDS("ALL_annotated.RDS")
# split <- SplitObject(ALL, split.by = "class")
for (subset in names(split)){
  message(paste0("Saving ", subset, " subset..."))
  print(split[[subset]])
  saveRDS(split[[subset]], file = paste0("class_annotated/",subset,".RDS"))
}
```

# Calculate each cell class   
  
Note that the Neuron and Unknown object were skipped becuase of their sizes
```{r}

# loop over files in folder
# 
SOs <- list.files('class_annotated', pattern='*\\.RDS', recursive=FALSE, full.names=TRUE)

for (file in SOs[11:12]){ # use this to run all except neuron and unknown SOs[-c(9, 13)]
  SO <- readRDS(file)
  print(file)
  print(SO)
  # SO <- dropPass(SO)
  SO <- calcDrop(SO)
  message(paste0("Writing ", basename(file), " to disk..."))
  saveRDS(SO, paste0("CLASS/", basename(file)))
}
```



# ALL PASSED CELLS AS A SINGLE OBJECT

## Split All cells into a only passed object
```{r}

QCsplit <- SplitObject(ALL, split.by = "reason")
```
```{r}
saveRDS(QCsplit[["passed"]], file = "ALL/ALL.RDS")
```



## Calculate the ALL, NEURON and UNKNOWN object  
  
Note that in order to compute the SCTransform for the NEURON and OBJECT object, the conserve.memory option is set to TRUE. This resulting that only variable genes are returned

```{r}
NEURON <-readRDS('class_annotated/NEURON.RDS')

```

```{r}

  # NEURON <- dropPass(NEURON)
  #stopifnot(length(NEURON@meta.data$orig.ident) >100)
  message("Performing Normalization and Variance Stablization")
  NEURON <- suppressWarnings(SCTransform(NEURON,
                                     vars.to.regress = "percent.mito" ,
                                     verbose = TRUE,
                                     return.only.var.genes = TRUE,
                                     conserve.memory = TRUE
                                      ))
  
  message("Running PCA")

  NEURON <- suppressWarnings(RunPCA(NEURON, verbose = FALSE))
  # write test to extract max number of PCs
  # set dims 
  message("Testing Dimensionality")
  # NEURON <- JackStraw(NEURON, num.replicate = 100)
  # NEURON <- ScoreJackStraw(NEURON, dims = 1:50)
  message("Running Dimension Reduction")
  NEURON <- RunUMAP(NEURON, dims = 1:30, verbose = FALSE)
  NEURON <- RunTSNE(NEURON, dims = 1:30, verbose = FALSE)
  # clustering
  message("Finding Nearest Neighbors")
  NEURON <- FindNeighbors(NEURON, dims = 1:30, verbose = FALSE)
  message("Finding Clusters")
  NEURON <- FindClusters(NEURON, verbose = FALSE)
  saveRDS(NEURON, "CLASS/NEURON.RDS")
```

```{r}
UNKNOWN <-readRDS('class_annotated/UNKNOWN.RDS')
```

```{r}
 # UNKNOWN <- dropPass(UNKNOWN)
  #stopifnot(length(UNKNOWN@meta.data$orig.ident) >100)
  message("Performing Normalization and Variance Stablization")
  UNKNOWN <- suppressWarnings(SCTransform(UNKNOWN,
                                     vars.to.regress = "percent.mito" ,
                                     verbose = TRUE,
                                     return.only.var.genes = TRUE,
                                     conserve.memory = TRUE
                                      ))
  
  message("Running PCA")

  UNKNOWN <- suppressWarnings(RunPCA(UNKNOWN, verbose = FALSE))
  # write test to extract max number of PCs
  # set dims 
  message("Testing Dimensionality")
  # UNKNOWN <- JackStraw(UNKNOWN, num.replicate = 100)
  # UNKNOWN <- ScoreJackStraw(UNKNOWN, dims = 1:50)
  message("Running Dimension Reduction")
  UNKNOWN <- RunUMAP(UNKNOWN, dims = 1:30, verbose = FALSE)
  UNKNOWN <- RunTSNE(UNKNOWN, dims = 1:30, verbose = FALSE)
  # clustering
  message("Finding Nearest Neighbors")
  UNKNOWN <- FindNeighbors(UNKNOWN, dims = 1:30, verbose = FALSE)
  message("Finding Clusters")
  UNKNOWN <- FindClusters(UNKNOWN, verbose = FALSE)
  saveRDS(UNKNOWN, "CLASS/UNKNOWN.RDS")
```

```{r}
ALL <-readRDS('ALL/ALL.RDS')
```

```{r}
 # ALL <- dropPass(ALL)
  #stopifnot(length(ALL@meta.data$orig.ident) >100)
  message("Performing Normalization and Variance Stablization")
  ALL <- suppressWarnings(SCTransform(ALL,
                                     vars.to.regress = "percent.mito" ,
                                     verbose = TRUE,
                                     return.only.var.genes = TRUE,
                                     conserve.memory = TRUE
                                      ))
  
  message("Running PCA")

  ALL <- suppressWarnings(RunPCA(ALL, verbose = FALSE))
  # write test to extract max number of PCs
  # set dims 
  message("Testing Dimensionality")
  # ALL <- JackStraw(ALL, num.replicate = 100)
  # ALL <- ScoreJackStraw(ALL, dims = 1:50)
  message("Running Dimension Reduction")
  ALL <- RunUMAP(ALL, dims = 1:30, verbose = FALSE)
  ALL <- RunTSNE(ALL, dims = 1:30, verbose = FALSE)
  # clustering
  message("Finding Nearest Neighbors")
  ALL <- FindNeighbors(ALL, dims = 1:30, verbose = FALSE)
  message("Finding Clusters")
  ALL <- FindClusters(ALL, verbose = FALSE)
  saveRDS(ALL, "ALL/ALL_calc.RDS")
```

# Will replace previous sections for the large SOs
```{r}
bigSOs <- c('class_annotated/NEURON.RDS', 'class_annotated/UNKNOWN.RDS', 'ALL/ALL.RDS')
```
```{r}
# TODO: Finish this cell by adding saveRDS to correct place
for (file in bigSOs){
  print(file)
  SO <- readRDS(file)
  message("Performing Normalization and Variance Stablization")
  SO <- suppressWarnings(SCTransform(SO,
                                     vars.to.regress = "percent.mito" ,
                                     verbose = TRUE,
                                     return.only.var.genes = TRUE,
                                     conserve.memory = TRUE
                                      ))
  
  message("Running PCA")

  SO <- suppressWarnings(RunPCA(SO, verbose = FALSE))
  # write test to extract max number of PCs
  # set dims 
  message("Testing Dimensionality")
  # SO <- JackStraw(SO, num.replicate = 100)
  # SO <- ScoreJackStraw(SO, dims = 1:50)
  message("Running Dimension Reduction")
  SO <- RunUMAP(SO, dims = 1:30, verbose = FALSE)
  SO <- RunTSNE(SO, dims = 1:30, verbose = FALSE)
  # clustering
  message("Finding Nearest Neighbors")
  SO <- FindNeighbors(SO, dims = 1:30, verbose = FALSE)
  message("Finding Clusters")
  SO <- FindClusters(SO, verbose = FALSE)
  saveRDS(SO, "ALL/ALL_calc.RDS")
}
```

